---
title: "Global Burden of Disease - West Sussex 2019 summary"
author: "Dr Rich Tyler"
date: '2021'
output: pdf_document
---

```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = '~/GitHub/gbd_2019/')
```


```{r data, echo=FALSE}
library(easypackages)

libraries(c("readxl", "readr", "plyr", "dplyr", "ggplot2", "png", "tidyverse", "reshape2", "scales", "viridis", "rgdal", "officer", "flextable", "tmaptools", "lemon", "fingertipsR", "jsonlite"))

data_directory <- './Source_files/Raw'
output_directory <- './Outputs'
meta_directory <- './Source_files'

codebook <- read_csv(paste0(meta_directory,'/IHME_GBD_2019_CODEBOOK_Y2020M11D25.csv'))
cause_hierarchy <- read_excel(paste0(meta_directory, '/IHME_GBD_2019_CAUSE_HIERARCHY_Y2020M11D25.xlsx'))

le_wsx <- fromJSON(paste0(output_directory, '/le_wsx.json')) %>% 
  mutate(Sex = factor(Sex, levels = c('Male', 'Both', 'Female')))

le_wsx_gap <- le_wsx %>% 
  select(!Sub_optimal_health) %>% 
  pivot_longer(names_to = 'Measure',
               cols = c('LE', 'HALE')) %>% 
  select(Name, Sex, Year, Measure, value) %>% 
  pivot_wider(names_from = 'Sex',
              values_from = 'value') %>% 
  mutate(Sex_gap_female_minus_male = Female - Male) %>% 
  select(Name, Year, Measure, Sex_gap_female_minus_male) %>% 
  pivot_wider(names_from = 'Measure',
              values_from = 'Sex_gap_female_minus_male')

wsx_compare <- fromJSON(paste0(output_directory, '/wsx_compare_df.json'))

wsx_ranks_compare <- fromJSON(paste0(output_directory, '/wsx_ranks_df.json'))

wsx_ranks_compare %>% 
  filter(sex == 'Both') %>% 
  mutate(Label = paste0(cause, ' ', format(round(`DALYs (Disability-Adjusted Life Years)_value`,1), big.mark = ',', trim = TRUE), ' DALYs (', format(round(Deaths_value, 0), big.mark = ',', trim = TRUE), ' deaths)')) %>% 
  filter(metric == 'Number') %>% 
  arrange(`DALYs (Disability-Adjusted Life Years)_rank`) %>% 
  view()

le_comp <- read_csv('https://raw.githubusercontent.com/psychty/gbd_2019/ecd33659ad73f41d90e9296cc3d55da3555e420d/Source_files/Raw/Life_expectancy_SE_2019.csv') %>% 
  filter(location %in% c('West Sussex','England','South East England')) %>% 
  filter(year == 2019) %>% 
  # filter(sex != 'Both') %>% 
  filter(age == '<1 year') %>% 
  select(sex, location, val) %>% 
  mutate(val = round(val, 1)) %>% 
  pivot_wider(values_from = val,
              names_from = location)

hale_comp <- read_csv('https://raw.githubusercontent.com/psychty/gbd_2019/ecd33659ad73f41d90e9296cc3d55da3555e420d/Source_files/Raw/Health_Adjusted_Life_expectancy_SE_2019.csv') %>% 
  filter(location %in% c('West Sussex','England','South East England')) %>% 
  filter(year == 2019) %>% 
  # filter(sex != 'Both') %>% 
  filter(age == '<1 year') %>% 
  select(sex, location, val) %>% 
  mutate(val = round(val, 1)) %>% 
  pivot_wider(values_from = val,
              names_from = location)

population_df <- fromJSON(paste0(output_directory, '/wsx_population.json'))

```

## Key findings

``` {r, infographic echo = FALSE}

wsx_compare %>% 
  filter(Year == 2019) %>% 
  filter(Sex == 'Both') %>% 
  filter(Cause == 'All causes') %>% 
  # filter(Cause %in% c('All causes', 'Communicable, maternal, neonatal, and nutritional diseases', 'Non-communicable diseases','Injuries')) %>% 
  view()

```

``` {r, infographic_top_five echo = FALSE}

wsx_compare %>% 
  filter(Year == 2019) %>% 
  filter(Sex == 'Both') %>% 
  filter(Measure == 'DALYs (Disability-Adjusted Life Years)') %>% 
  filter(!Cause %in% c('All causes', 'Communicable, maternal, neonatal, and nutritional diseases', 'Non-communicable diseases','Injuries')) %>% 
  view()

```

# Measures of burden, what is the Global Burden of Disease?

The Global Burden of Disease (GBD) is one of the most comprehensive, worldwide, epidemiological collaborations describing life expectancy, as well as mortality and morbidity from major diseases and injuries as well as risk factors at local, regional, national, and global levels.

## Exploring the data

## Is life expectancy improving in West Sussex?

Life expectancy increased from 1990 to 

``` {r Life_expectancy, echo = FALSE} 

ggplot(le_wsx,
       aes(x = Year,
           colour = Sex,
           fill = Sex)) +
  geom_line(aes(y = LE)) +
  geom_point(aes(y = LE)) +
  geom_line(aes(y = HALE)) +
  labs(title = 'Life expectancy and healthy life expectancy at birth by sex; West Sussex; 1990-2019',
       y = 'Years',
       x = 'Year') +
  scale_y_continuous(limits = c(0,90)) +
  scale_x_continuous(limits = c(1990, 2019),
                     breaks = seq(1990, 2019, 1)) +
  scale_color_manual(values = c('#fd6400', '#172243', '#00c3ff')) +
  scale_fill_manual(values = c('#fd6400', '#172243', '#00c3ff')) +
  theme(legend.position = 'none')

```
